Comparing ChatGPT-4 to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6942 Deaths in Sierra Leone from 2019-2022

### 2024-01-24 manuscript/manuscript.tex
Working on reproducibility

Initially, 11,920 records from the HEAL-SL study \cite{njalauniversityHealthySierraLeone2023,carshon-marshChildMaternalAdult2022} were collected from dual-coded EVA, where each record was randomly coded by two different physicians that assigned CODs as International Classification of Diseases Revision 10 (ICD-10) codes \cite{worldhealthorganizationICD10InternationalStatistical2011}. Physicians were able to assign CODs for 11,820 of the 11,920 records, where 100 of these records could not be assigned a COD due to missing or inadequate information (e.g. low quality narrative, data loss). To determine if two codes were in physician agreement per record, codes were compared for similarity using Central Medical Evaluation Agreement 10 (CMEA-10) codes, which groups a range of similar ICD-10 codes together indicating codes in agreement \cite{aleksandrowiczPerformanceCriteriaVerbal2014} (see Additional File 1). When codes were not in agreement, a record enters the reconciliation phase, where the two physicians were provided reasoning and initial codes from each other to: (1) keep their initial code (2) assign the other physician's code or (3) assign a new code. If codes were not in agreement after the reconciliation phase, a record enters the adjudication phase, where a third senior physician evaluates both physicians' reasoning and codes before and after reconciliation, and assigns a final code based on their evaluation. The 11,820 physician coded records were further filtered for records where both physicians agreed on the assigned codes (records that were not reconciled or adjudicated), resulting in 6942 physician agreed records, where . Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \cite{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}.

@@

For InterVA-5 and InSilicoVA, the standardized questionnaire data from the HEAL-SL EVA were first converted into 2016 World Health Organization (WHO) VA questionnaire revision 1.5.1 
\cite{} Open Data Kit (ODK) format \cite{worldhealthorganizationODKVerbalAutopsy2022,nafundiODKCollectData2023}, followed by further conversion into OpenVA format \cite{liOpenVAToolkitVerbal2023} using the \verb|pyCrossVA| Python package \cite{thomasPycrossvaPrepareData}, and used as input for InterVA-5 and InSilicoVA. The \verb|Nsim| (number of iterations to run) parameter for InSilicoVA was set to 10,000, while the \verb|HIV| (level of prevalence of human immunodeficiency virus) and \verb|Malaria| (level of prevalence of Malaria) parameters were both set to \verb|'h'| (high).

@@

Of the 6942 physician agreed records, ChatGPT-3.5 and ChatGPT-4 were able to assign CODs for 6932 and 6912 records respectively, while InterVA-5 and InSilicoVA both assigned CODs for 6833 records. All 6833 records with WHO VA 2016 v1.5 codes \cite{worldhealthorganizationVerbalAutopsyStandards2016} from InterVA-5 and InSilicoVA output were converted into ICD-10 codes. After all model outputs contained ICD-10 codes, they were further converted to CGHR-10 codes that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 years or older), child (28 days to 11 years), and neonatal (under 28 days) age groups (see Additional file 2). The 6932 and 6912 records with ICD-10 codes from ChatGPT-3.5 and ChatGPT-4 output were converted into CGHR-10 codes, resulting in 6894 and 6857 records, where 38 and 55 records did not have matching CGHR-10 codes respectively. The 6833 records with ICD-10 codes from InterVA-5 and InSilicoVA output were also converted into CGHR-10 codes, resulting in and 6805 and 6729 records, where 28 and 104 records did not have matching CGHR-10 codes respectively. The 6942 physician agreed records with COD assignments in ICD-10 were also converted to CGHR-10 codes, which resulted in 6939 physician agreed records for comparison across physician and model assignments (three records in ICD-10 were not captured by the CGHR-10 ranges).

@@

>>>>

Initially, 11,920 records from the HEAL-SL study \cite{njalauniversityHealthySierraLeone2023,carshon-marshChildMaternalAdult2022} were collected from dual-coded EVA, where each record was randomly coded by two different physicians that assigned CODs as International Classification of Diseases Revision 10 (ICD-10) codes \cite{worldhealthorganizationICD10InternationalStatistical2011}. Physicians were able to assign CODs for 11,820 of the 11,920 records, where 100 of these records could not be assigned a COD due to missing or inadequate information (e.g. low quality narrative, data loss). To determine if two codes were in physician agreement per record, codes were compared for similarity using Central Medical Evaluation Agreement 10 (CMEA-10) codes, which groups a range of similar ICD-10 codes together indicating codes in agreement \cite{aleksandrowiczPerformanceCriteriaVerbal2014} (see Additional File 1). When codes were not in agreement, a record enters the reconciliation phase, where the two physicians were provided reasoning and initial codes from each other to: (1) keep their initial code (2) assign the other physician's code or (3) assign a new code. If codes were not in agreement after the reconciliation phase, a record enters the adjudication phase, where a third senior physician evaluates both physicians' reasoning and codes before and after reconciliation, and assigns a final code based on their evaluation. The 11,820 physician coded records were further filtered for records where both physicians agreed on the assigned codes (records that were not reconciled or adjudicated), resulting in 6942 physician agreed records, containing 3829 adult, 2636 child, and 477 neonatal records. Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \cite{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}.

@@

For InterVA-5 and InSilicoVA, the standardized questionnaire data from the HEAL-SL EVA were first converted into 2016 World Health Organization (WHO) VA questionnaire revision 1.5.1 
\cite{dipasqualeReleaseODK20162016} Open Data Kit (ODK) format \cite{worldhealthorganizationODKVerbalAutopsy2022,nafundiODKCollectData2023}, followed by further conversion into OpenVA format \cite{liOpenVAToolkitVerbal2023} using the \verb|pyCrossVA| Python package \cite{thomasPycrossvaPrepareData}, and used as input for InterVA-5 and InSilicoVA. The \verb|Nsim| (number of iterations to run) parameter for InSilicoVA was set to 10,000, while the \verb|HIV| (level of prevalence of human immunodeficiency virus) and \verb|Malaria| (level of prevalence of Malaria) parameters were both set to \verb|'h'| (high).

@@

Of the 6942 physician agreed records, ChatGPT-3.5 and ChatGPT-4 were able to assign CODs for 6932 and 6912 records respectively, while InterVA-5 and InSilicoVA both assigned CODs for 6833 records. All 6833 records with WHO VA 2016 v1.5 codes \cite{worldhealthorganizationVerbalAutopsyStandards2016} from InterVA-5 and InSilicoVA output were converted into ICD-10 codes. After all model outputs contained ICD-10 codes, they were further converted to CGHR-10 codes that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 years or older), child (28 days to 11 years), and neonatal (under 28 days) age groups (see Additional file 2). The 6932 and 6912 records with ICD-10 codes from ChatGPT-3.5 and ChatGPT-4 output were converted into CGHR-10 codes, resulting in 6894 and 6857 records, where 38 and 55 records did not have matching CGHR-10 codes respectively. The 6833 records with ICD-10 codes from InterVA-5 and InSilicoVA output were also converted into CGHR-10 codes, resulting in and 6805 and 6729 records, where 28 and 104 records did not have matching CGHR-10 codes respectively. The 6942 physician agreed records with COD assignments in ICD-10 were also converted to CGHR-10 codes, which resulted in 6939 physician agreed records for comparison across physician and model assignments (three records in ICD-10 were not captured by the CGHR-10 ranges). The 6939 records contained 3826 adult, 2636 child, and 477 neonatal records.

@@

Dimensions and cost of LLM versus physician.

Reliance on third-party hosting.
